Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa

### 2024-04-04 paper/paper.qmd
updated paper, data cleaning

---
title: "Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa"
author: 
  - Catherine Punnoose
thanks: "Code and data are available at: https://github.com/catherinee1216/AutoTheft_Ottawa.git."
date: today
date-format: long
abstract: "The surge in auto theft rates across Canada has prompted car owners to become more vigilant regarding their vehicle's security. This paper gives a thorough examination of auto theft occurrences within Canada's capital, Ottawa. Focusing on the years 2018-2022, the paper provides a deep dive into the types of vehicles brands, model, and color targeted and offers insights into the neighborhoods within Ottawa exhibiting elevated rates of auto theft. Additionally, the study assesses factors such as the recovery rate of stolen cars, and provides solutions to improve vehicle security by understanding auto theft dynamics in the region."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(janitor)
library(kableExtra)
library(maps)
```

@@

@@

auto_Data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))

@@

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

auto_Data |>
  ggplot(aes(x = theft_year)) +
  geom_histogram(stat = "count")
```

@@

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

>>>>

---
title: "Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa"
author: 
  - Catherine Punnoose
thanks: "Code and data are available at: https://github.com/catherinee1216/AutoTheft_Ottawa.git."
date: today
date-format: long
abstract: "The surge in auto theft rates across Canada has prompted car owners to become more vigilant regarding their vehicle's security. This paper gives a thorough examination of auto theft occurrences within Canada's capital, Ottawa. Focusing on the years 2018-2022, the paper provides a deep dive into the types of vehicles brands, model, and color targeted and offers insights into the neighborhoods within Ottawa exhibiting elevated rates of auto theft. Additionally, the study assesses factors such as the recovery rate of stolen cars, and provides solutions to improve vehicle security by understanding auto theft dynamics in the region."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

library(tidyverse)
library(janitor)
library(kableExtra)
library(maps)
library(ggmap)
library(arrow)

auto_Data <- read_parquet(here::here("data/analysis_data/analysis_data.parquet"))
```

@@

[TO BE EDITED]

@@

The data was cleaned by using the open source statistical programming language R [@citeR]...[TO UPDATE]

```{r}
#| label: theft-year
#| fig-cap: "Average Auto Theft Crime Reported in ottawa from 2018-2023"
#| echo: false
#| warning: false
#| message: false

auto_Data |>
  ggplot(aes(x = theft_year)) +
  geom_histogram(stat = "count")
```
```{r}
#| label: theft-brand
#| fig-cap: "Number of Cars Stolen by Car Brand in 2018"
#| echo: false
#| warning: false
#| message: false

brand_theft_2018 <- auto_Data |>
  filter(theft_year == "2018")

brand_theft_2018 |>
  ggplot(aes(x = vehicle_make, fill = vehicle_make)) +
  geom_histogram(stat = "count")

@@

```{r}
#| echo: false
#| warning: false
#| message: false

# creating map from section 5.4.1 in the textbook

# Define the bounding box for Ottawa
bbox <- c(left = -75.9, bottom = 45.3, right = -75.5, top = 45.5)

# Get the map of Ottawa using get_stadiamap function
ottawa_stadia_map <- get_stadiamap(bbox, zoom = 11, maptype = "stamen_toner_lite")

# Plot the map using ggmap
ggmap(ottawa_stadia_map) # map prints

# ---- not working -----------
#filtering dataset to get auto theft reports per neighborhood in 2018

#neighborhood_theft_2018 <- auto_Data |>
#  filter(theft_year == "2018") |>
#  select(neighbourhood, x, y)

#print(neighborhood_theft_2018)

# map does not print with elements plotted

#ggmap(ottawa_stadia_map, extent = "normal", maprange = FALSE) +
#  geom_point(data = neighborhood_theft_2018,
#             aes(x = 45.3, y = 45.5, colour = neighbourhood),
#             alpha = 0.7) +
#  scale_color_brewer(name = "2019 Division", palette = "Set1") +
#  labs(x = "Longitude",
#       y = "Latitude") +
#  theme_minimal() +
#  theme(panel.grid.major = element_blank(),
#        panel.grid.minor = element_blank())

@@

## Possible Reasons for Increasing Auto Theft Rates

## Methods to Increase Vehicle Security

## Auto Theft Rates Projection
